---
title: "Math250 HW6 Inference, etc"
author: "Ilana Berlin" 
date: "November 5th" 
format: html
embed-resources: true
code-fold: true
---

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(ISLR2)
```

### Problem 1 
> The first two problems in this set refer to the ISLR2::Carseats set, which includes simulated sample data for 400 stores selling a particular brand of child carseat. We view this data as a simple random sample from a larger population. Does this data provide evidence against the company’s claim that the average price of their own seats is $110? Follow all best practices from class, including hypotheses, a relevant visualization, and a conclusion in non-technical language.

Null hypothesis: $H: \mu = 100$

Alternative: $H_a: \mu \ne 100$ 

```{r}
#| label: prices histogram 
#| fig-alt: a histogram showing the distribution of car seat prices

Carseats %>% 
  ggplot(aes(x = Price)) +
  geom_histogram(bins = 25) + 
  geom_vline(xintercept = 100, 
             color = "red") + 
  theme_minimal()
```
```{r}
#| label: prices t test
Carseats$Price %>% 
  t.test(mu = 100)
```
The average price of car seats is not $100 because if it was then it would be incredibly rare to get data like ours. 

### Problem 2
> Compute a level 90% confidence interval for the mean competitor price in the population from which this sample is drawn. What is the sample mean? What is the margin of error?

```{r}
#| label: competitor price

Carseats$CompPrice %>% 
  t.test(conf.level = .90)

```
The sample mean is \$124.98 $\pm$ 1.2591. 

### Problem 3
> Watch my video on contingency tables. Use these tools to create a table with counts of penguin species and home islands in the penguins set. Include marginal counts in your table.

```{r}
#| label: contigency table

t <- table(penguins$species, 
      penguins$island)
addmargins(t)
```


### Problem 4
> Push your solutions to this set to your GitHub page, including both the Quarto source and the rendered html. Use file names of the form, “math250_nov5_yourname.” Provide a clickable link to the repository on your personal GitHub where these files can be found.

[Ilana Berlin Github - Math250 Sandbox](https://github.com/IlanaB11/Math250_GithubSandbox)

