---
title: "Chi Squared HW"
format: html
embed-resources: true
code-fold: true
author: Ilana Berlin
date: Nov 10th 2025
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(readxl)

at_small <-read.csv("attrition_small.csv")
```

### Problem 1

> The company claims that 10% of employees report their work/life balance as “best,” 65% report it as “better,” 20% as “good,” and only 5% as “bad.” Is this claim reasonable in light of the sample data?

$H_o$: The proportions stated by the company are accurate

$H_a$: The proportions stated by the company are inaccurate

```{r}
#| label: work life balance visual
#| fig-alt: a bar chart showing the work life balance of employees

at_small %>% 
  ggplot(aes(x = WorkLifeBalance)) +
  geom_bar() + 
  theme_minimal() +
  labs(x = "Work Life Balance", 
       y = "Count", 
       title = "Work Life Balance of 250 Employees")

```
```{r}
#| label: work life chi squared

obs <- table(at_small$WorkLifeBalance)
h_null <- c(0.05, 0.1, 0.65, 0.2 )

chisq.test(obs, 
           p = h_null)
```
If the proportions that the company claims are true than it would be reasonable to get sample data like ours. 

### Problem 2

> The company claims that the education field of 55% of their employees is in the life science, 20% is in marketing, 20% is in technical, and 5% is other. Is this claim reasonable in light of the sample data?

$H_o$: The proportions stated by the company are accurate

$H_a$: The proportions stated by the company are inaccurate

```{r}
#| label:  edu visual
#| fig-alt: a bar chart showing the education field of employees

at_small %>% 
  ggplot(aes(x = EducationField)) +
  geom_bar() + 
  theme_minimal() +
  labs(x = "Education Field", 
       y = "Count", 
       title = "Education Field of 250 Employees")

```
```{r}
#| label: edu chi squared

obs_edu <- table(at_small$EducationField)
h_null_edu <- c(0.55, 0.20, 0.05, 0.2)

chisq.test(obs_edu, 
           p = h_null_edu)

```
If the proportions stated by the company are true that it would be unusual to get sample data like ours. 
